---
title: "R Notebook"
author: "Tom Richebacher"
date: "7/11/2023"
output: html_notebook
---

```{r call-libraries, echo=TRUE, warning=TRUE}

R_LIBS_USER <- "C:/Users/Tom/OneDrive/R_code/win-library/4.2"

library(formatR)
library(HiClimR)
library(lintr)
library(data.table)
library(caret)
library(ggplot2)
library(MLmetrics)
library(janitor)
library(partykit)
library(stringr)
library(DataExplorer)
library(smbinning)
library(funModeling)
library(SmartEDA)
library(fastDummies)
library(rattle)
library(rpart.plot)
library(RColorBrewer)
library(treeClust)

library(doParallel)
detectCores()

registerDoParallel(cores = 8)


```

## [Initialize global variables]{.underline}

```{r initialize-global-variables, warning=TRUE}
options(scipen = 999)
use.integer64 <- FALSE

#"REG" = regression, "B_CLASS" = binary class, "M_CLASS" = multiclass
analysis_type <- "B_CLASS"
input_files <- "1"
project <- "Credit"
report_path <- "C:/Users/trich/OneDrive/R_code/rtest/R_testing"
data_path <- "C:/Users/trich/OneDrive/R_code/credit"
set.seed(4193)
```

# [Get Files:]{.underline}

## Read data

```{r message=TRUE, warning=TRUE}
read_data <- function(filename) {
       fread(filename,
                # Set separator as comma
                sep = ",",
                # Read all rows - use nrows = xxx to read partial file
                nrows = -1,
                # Check column names for validity
                check.names = TRUE,
                #read random sample
                #n = 50000,
                # Set header as first row
                header = TRUE,
                # treat all integer columns as 64-bit integers
                #integer64 = 'numeric',
                #Use quote to specify character used to quote strings in the file
                #quote = "'",
                #specify the possible string representations of missing values
                na.strings = c("NA", "N/A", "null", "", "unknown"))
 }

# Call function to read in data files
if (input_files == "1") {
  DT <- read_data(paste0(data_path, "/DT.csv"))
  saveRDS(DT, paste0(data_path, "/DT.Rds"))
} else if (input_files == "2") {
  train <- read_data("train.csv")
  train[, file := "train"]
  saveRDS(train, paste0(data_path, "/train.Rds"))
  test <- read_data("test.csv")
  test[, file := "test"]
  saveRDS(test, paste0(data_path, "/test.Rds"))
}

rm(read_data)

#for speed testind a data dublication function
dt_dup <- rbindlist(rep(list(DT), 10000))
DT <- copy(dt_dup)
rm(dt_dup)
start_time <- Sys.time()
```

## Stack train and test file

```{r eval=FALSE, include=FALSE}
if (input_files == 2) {

    # function to identify column header differences
    get_feature_names <- function(file1, file2) {
        # find the names that are common between the tables
        var_names <- intersect(names(file1), names(file2))
        # find columns different in file1
        var_names_file1 <- setdiff(names(file1), names(file2))
        # find columns different in file 2
        var_names_file2 <- setdiff(names(file2), names(file1))
        # show all differences between the two files
        col_diff <- paste0(var_names_file1, var_names_file2)
        return(col_diff)
    }

    # return column header difference
    get_feature_names(train, test)

    # #assign Id if needed setnames(train, 'train_id', 'Id') setnames(test, 'test_id', 'Id')

    # stack train and test
    DT <- rbind(train, test, fill = TRUE, use.names = TRUE)
    rm(train, test, get_feature_names)
}

```

###### Join Files

```{r eval=FALSE, include=FALSE}

DT <-  merge(DT, people, by = "people_id", all.x = TRUE)
DT$Id <- seq.int(nrow(DT))
 
rm(people)
```

###### Sample File

```{r Sample-file, eval=FALSE, include=FALSE}
DT_sample <- DT[sample(.N, 30000)]
```

## Create record Id

```{r}
#create index
DT[, Id := as.integer(.I)]

#move index to 1st column
setcolorder(DT, c("Id", colnames(DT)[1:(ncol(DT) - 1)]))
```

# [Column Work]{.underline}

## Clean column headers

```{r eval=FALSE, include=FALSE}
# identify non standard column headers
# Define a regex pattern for non-standard column names

non_standard_pattern <- "^[^A-Za-z_].*$"

# Identify non-standard column names
non_standard_cols <- names(DT)[grepl(non_standard_pattern, names(DT))]

# Print the non-standard column names
print(non_standard_cols)

# Clean column names
setnames(DT, clean_names(names(DT)))

setnames(DT, old = c("a", "d"), new = c("anew", "dnew"))

rm(non_standard_pattern, non_standard_cols)
```

## Create sample for DIST report if records \> 50,000

```{r}
if (nrow(DT) > 100000) {
   DT_dist_sample = DT[sample(.N, 50000)]
} else {
   DT_dist_sample <- DT
}
```

## Distribution Report

```{r paged.print=TRUE}
# Table with basic statistics
smbinning.eda(DT_dist_sample, rounding = 3)$eda

rm(DT_dist_sample)
```

## Change Column Classes

```{r eval=FALSE, include=FALSE}
#individual var
DT[, target := as.factor(target)]

#by specific vars
change_col_class <- c("return_dollar", "product_id")

#by column text spring
change_col_class <- grep("(cat|bin)", names(DT), value = TRUE)

#by all vars of a specific class
change_col_class <- colnames(DT[, .SD, .SDcols = sapply(DT, is.numeric)])

#identify binaries to convert
change_col_class <- names(Filter(function(x)uniqueN(na.omit(x)) <= 2 & max(x) == 1, DT))

#********************************************************************************#
#do actual conversion
DT[, (change_col_class) := lapply(.SD, as.numeric), .SDcols = change_col_class]
#********************************************************************************#
#*
rm(change_col_class)
```

## Delete unnecessary columns

```{r eval=FALSE, include=FALSE}

#delete columns with one value
rm_cols <- colnames(DT)[sapply(DT, function(x) length(unique(x))) == 1]

#delete specific cols
rm_cols <- c("V1", "id")

if (length(rm_cols) > 0) {
  DT[, (rm_cols) := NULL]
}

rm(rm_cols)
```

# [Row Work]{.underline}

## Get % missing

```{r eval=FALSE, include=FALSE}
# Calculate the number of missing values in each row
DT[, row_miss_orig := rowSums(is.na(DT))]

#run frequency count
freq(DT$row_miss_orig)

# filter out records if needed
#dt[x <= 3]
```

###### Omit rows where either vars 'x' or 'y' have NA

```{r eval=FALSE, include=FALSE}
# select columns
omit_row <- c("x", "y")

# see how many records are left
DT[complete.cases(DT[, ..omit_row]), ]
```

# Target Prep

## Target Analysis & Prep

```{r}
# assign target name
setnames(DT, "default", "target")

# move target to 1st column
setcolorder(DT, c("target", setdiff(names(DT), "target")))

# target distribution
if (analysis_type != "REG") {
    DT[, .(.N), by = .(target)]
    plot_bar(DT$target)
} else {
    plot_histogram(DT$target)
}

# rename target value when binary
if (analysis_type == "B_CLASS") {
DT[, target := ifelse(target == 1, "HaveNot", ifelse(target == 2, "Have", NA))]
}

# remove missing target records
DT[!is.na(target)]
```

###### Normalize Target Variable

```{r eval=FALSE, include=FALSE}
DT[, target := log(target + 1)]
```

## Append identifier to original column headers

```{r}
# get names to exclude
exclude_columns <- c("target", "Id")

# Get original data columnns
orig_columns <- setdiff(colnames(DT), exclude_columns)

# define the suffix
suffix <- "_orig"

# Append the suffix to the selected columns
setnames(DT, old = orig_columns, new = paste0(orig_columns, suffix))

rm(orig_columns, exclude_columns, suffix)
```

# Create table with data that bypasses the analysis

```{r include=FALSE}
# identify vars for bypass table
mov_var <- c("target", "Id")

# create bypass table
DT_report <- DT[, .SD, .SDcols = mov_var]

# save bypass table
saveRDS(DT_report, paste0(data_path, "/DT_tar.Rds"))

# remove bypass table unless needed
rm(mov_var)
gc()
```

# Data Exploration

## Create report sample if records \> 50,000

```{r}
if (nrow(DT) > 100000) {
   DT_report_sample = DT[sample(.N, 50000)]
} else {
   DT_report_sample <- DT
}
```

## General data profile report

```{r include=FALSE}

# create in depth data report
data_profile_report <- create_report(DT_report_sample, y = "target", report_title = paste0(project, "Data Profile Report"), output_file = paste0(project, "Profile Report"), output_dir = report_path)
```

## Summary statistics for char and cat vars to target

```{r include=FALSE}
# analysis of categorical data
stats_report_char_and_cat <- ExpCatStat(DT_report_sample[, -c("Id")],Target = "target",
                                  result = "Stat",
                                  clim = 10,        #max cat levels allowed
                                  nlim = 10,        #max unique values for num vars
                                  bins = 10,        #max number of bins
                                  Pclass = "Have",
                                  plot = TRUE,
                                  top = 20,
                                  Round = 2)
```

## Information value for char and cat vars to target

```{r include=FALSE}
info_report_char_and_cats <- ExpCatStat(DT_report_sample[, -c("Id")], Target = "target",
           result = "IV",
           clim = 5,          #max # of cats allowed
           nlim = 10,         # of levels for num vars
           bins = 10,         #max # of bins for cat & num vars
           Pclass = "Have")
```

## Summary stats for numeric data to target

```{r include=FALSE}
stats_report_num_vars <- ExpNumStat(DT_report_sample[, -c("Id")],
                                    by = "G",
                                    gp = "target",
                                    Qnt = NULL,
                                    Nlim = 10,
                                    MesofShape = 2,
                                    Outlier = TRUE,
                                    round = 2,
                                    weight = NULL,
                                    dcast = FALSE,
                                    val = NULL)
rm(DT_report_sample)
```

## Outlier analysis

```{r eval=FALSE, include=FALSE}
# #outlier analysis
# num_cols <- colnames(DT)[sapply(DT, is.numeric) & names(DT) != "Id"]
 
# outlier_analysis <- ExpOutliers(
# DT,
# varlist = num_cols,
# method = "boxplot",
# treatment = NULL,
# capping = c(0.05, 0.95),
# outflag = TRUE)
# test
```

## Missing values

Impute NA for incorrect missing value designation

```{r eval=FALSE, include=FALSE}
# replace values in all DT columns
DT[DT == 9999999999] <- NA_integer_

# replace values in specific columns
impute_miss_values <- c("a", "b")

DT[, impute_miss_values := lapply(.SD, function(x) replace(x, x == 9999999999, NA_integer_)), .SDcols = impute_miss_values]

rm(impute_miss_values)
```

Graph missing value

```{r}
plot_missing(DT) + theme_bw() + ggtitle("Percent Missing Values")
```

## High Cardinality

### Identify and collapse high cardinality vars

```{r}
library(Hmisc)
#get factor and char columns, exclude target & require > 2 values
char_fact_cols <- colnames(DT)[sapply(DT, function(x) {
  (is.character(x) || is.factor(x)) && length(unique(x)) > 2
}) & colnames(DT) != "target"]

# frequency distribution before combining
freq_before <- lapply(char_fact_cols, function(col) {
  freq_dt <- freq(DT[, .(value = get(col))], plot = FALSE, na.rm = TRUE)
  freq_dt <- cbind(freq_dt, column = col)
  return(freq_dt)
})

# creates table that shows freqs before transformation
freq_before <- rbindlist(freq_before)

#each value in a var needs to have at least 5% coverage 
threshold = 0.05   

#combine levels
if (length(char_fact_cols) > 0) {
  DT[, (char_fact_cols) := lapply(.SD, function(x) combine.levels(x, minlev = threshold)), .SDcols = char_fact_cols]
}

# after combine frequency distributions
freq_after <- lapply(char_fact_cols, function(col) {
  freq_dt <- freq(DT[, .(value = get(col))], plot = FALSE, na.rm = TRUE)
  freq_dt <- cbind(freq_dt, column = col)
  return(freq_dt)
})

# creates table with freqs after transformation
freq_after <- rbindlist(freq_after)

rm(threshold, char_fact_cols)
```

## Create binary variable from reduced cardinality

```{r}
# get factor and char columns & exclude relevant columns
conv_to_bins <- colnames(DT)[sapply(DT, function(x) {
  (is.character(x) || is.factor(x)) && length(unique(x)) > 2
}) & colnames(DT) != "target"]

# length of the initial DT data 
DT_cols <- ncol(DT) + 1

if (length(conv_to_bins) > 0) {
DT <- dummy_cols(
  DT,
  select_columns = conv_to_bins,
  remove_first_dummy = FALSE,
  remove_most_frequent_dummy = FALSE,
  ignore_na = FALSE,
  #remove_selected_columns = TRUE
  split = NULL)}

# Clean column names
clean_names(DT)
# Replace symbols with their corresponding abbreviations

#replace < with LT
setnames(DT, colnames(DT), str_replace_all(colnames(DT), "<", "LT"))

#replace > with GT
setnames(DT, colnames(DT), str_replace_all(colnames(DT), ">", "GT"))

#replace <= with LE
setnames(DT, colnames(DT), str_replace_all(colnames(DT), "<=", "LE"))

#replace => with GE
setnames(DT, colnames(DT), str_replace_all(colnames(DT), "=>", "GE"))

# Replace parentheses with an underscore
setnames(DT, colnames(DT), str_replace_all(colnames(DT), "\\(|\\)", "_"))

# Remove empty spaces from column names
setnames(DT, colnames(DT), str_replace_all(colnames(DT), "\\s+", ""))

#replace varies symbols with underscore
setnames(DT, colnames(DT), str_replace_all(colnames(DT), "[&|/|,]", "_"))

# Remove underscores at the end of column names
setnames(DT, colnames(DT), str_replace_all(colnames(DT), "_$", ""))

# define the suffix to name the bin columns
suffix <- "_bin"

# Get the bin column names to update
cols_to_update <- colnames(DT)[DT_cols:length(DT)]

# Append the suffix to the selected columns
setnames(DT, old = cols_to_update, new = paste0(cols_to_update, suffix))

# remove unnecessary orig_ identifier in the bin column
setnames(DT, colnames(DT), str_replace_all(colnames(DT), "orig_", ""))

#clean-up
rm(conv_to_bins, DT_cols, cols_to_update, suffix)
```

## Create bins based on non-activity

```{r eval=FALSE, include=FALSE}
#creates dummy rows to fill in missing rows based on all combinations of
#available character, factor, and date columns (if not otherwise specified).

dummy_rows(
.data,
select_columns = NULL,
dummy_value = NA,
dummy_indicator = FALSE
)
```

## Create weighted binaries

```{r}
# List of binary columns
weighted_bin_cols <- colnames(DT)[grep("_bin", colnames(DT))]

#create weighted binary
DT[, paste0(weighted_bin_cols, "_perc") := lapply(.SD, function(col) ifelse(col == 1, sum(col == 1) / nrow(DT), 0)), .SDcols = weighted_bin_cols]

rm(weighted_bin_cols)
```

# LAT/LONG work

```{r eval=FALSE, include=FALSE}
library(geosphere)

#Dist from centroid
DT$longmean <- mean((DT$longitude), na.rm = TRUE) / 1e6
DT$latmean  <- mean((DT$latitude), na.rm = TRUE) / 1e6

#Adjusted long lat
DT$longitude1 <- DT$longitude / 1e6
DT$latitude1 <- DT$latitude / 1e6

# Haversine distance
ncol(DT)
DT$geodist <- distHaversine(p[,31:32], DT[, 33:34])

rm_geo_cols <- c("longmean", "latmen", "longitude", "latitude1")

DT <- DT[(rm_geo_cols) := NULL]
```

# Feature Creation

## Row level - binary transformations

```{r eval=FALSE, include=FALSE}

#identify binary variables
create_bin_vars <- names(Filter(function(x) uniqueN(na.omit(x)) <= 2 & max(x) == 1, DT))

#create variable that sums the values in Bin_vars 
DT[, DT_sums  :=  rowSums(.SD, na.rm = TRUE), .SDcols = c(create_bin_vars)]

#compute mean of values in Bin_vars 
DT[, DT_means := rowMeans(.SD, na.rm = TRUE), .SDcols = c(create_bin_vars)]

#concatenate binary variable across columns and make it a factor variable
DT[, bin_pattern := do.call(paste0,.SD),.SDcols = create_bin_vars]
DT[, bin_pattern := as.factor(bin_pattern)]

# count zeros in row
DT[, num_cols_eq_0 := Reduce(sum, lapply(.SD, function(x) na.omit(x) == 0))]

#indicate if there are more than 4 missing values
DT[, high_nas := ifelse(amount_nas > 4, 1,0)] 

rm(Bin_vars)
```

### Row level aggregation by specific columns

```{r eval=FALSE, include=FALSE}

#conditionalby header identification aggregation 
#product_up <- names(DT_tar)[which(regexpr("M_UP_1", names(DT_tar)) > 0)]
#DT_tar[, prod_up := rowSums(.SD), by=ncodpers, .SDcols = product_up]  

cols_title <-  c("reg", "car", "calc")
#cols_title <- names(DT3)[which(regexpr(p, names(DT3)) > 0)]

 for (p in cols_title) 
  {
    cols_title <- names(DT3)[which(regexpr(p, names(DT3)) > 0)]   
    DT3[, paste0(p, '_sum')    := rowSums(.SD, na.rm = TRUE),
        .SDcols = c(cols_title)] 
    DT3[, paste0(p, '_mean')   := rowMeans(.SD, na.rm = TRUE),
        .SDcols = c(cols_title)]
    DT3[, paste0(p, '_ratio')  := ifelse(get(paste0(p, '_mean')) == 0,0,
                                             (get(paste0(p, '_sum'))) / get(paste0(p, '_mean'))) ]
  }

```

### Casting to wide data when multiple ID exist - like customer id

```{r eval=FALSE, include=FALSE}
#create new summary table with sum, mean, count


DT_summary_wide <- dcast(DT, customer_id ~ product_id + product_category,
                         value.var = c("revenue_dollar", "return_dollar"), 
                         fill = 0,
                         fun.aggregate = list(sum, mean, length))

#Create customer summary
DT_customer_summary <- DT[, .(dol_revenue = sum(revenue_dollar),
                              dol_returns = sum(return_dollar),
                              cnt_purchases = length(product_id),
                              cnt_prod_cats = .N,   
                              cnt_returns = .N),
                          by = .(customer_id)]

# Combine customer and order summaries
DT <- DT_summary_wide[DT_customer_summary, on = "Id"]

```

### Column level

```{r eval=FALSE, include=FALSE}
#create new variable in DT
DT[,`:=`(u = z + 1, v = z - 1)] 
```

# Create interaction vars

## Select column set used for interaction development

```{r}
#assign method to rpart and make sure categorical variable is factor
#DT[, target := as.factor(target)]

# Select original columns 
orig_cols <- c("target",colnames(DT)[grep("_orig", colnames(DT))])

# Select numeric columns with "_orig" in the column header
orig_num_cols <- colnames(DT)[sapply(DT, function(x) is.numeric(x) || is.integer(x)) & grepl("_orig", colnames(DT))]

# select new binary binary columns where "bin" is at the end of column name
new_bin_cols <- c("target", grep("bin$", colnames(DT), value = TRUE))

# select all weighted binary columns
new_bin_wgt_cols <- c("target",colnames(DT)[grep("_perc", colnames(DT))])

# assign chosen selection to analys_cols
analys_cols <- c(new_bin_wgt_cols, orig_cols)
```

Create modeling sample if records \> 50000

```{r}
if (nrow(DT) > 100000) {
   DT_model_sample = DT[sample(.N, 50000)]
} else {
   DT_model_sample <- DT
}
```

## Chaid Tree

```{r}

if (analysis_type == "REG") {
  method_m = "anova" 
  } else {
  method_m = "class"
}

#confusion matrix requires target is factor
DT_model_sample[, target := as.factor(target)]

#minimum oberservation in a terminal node is # of records * x
min_node_split = nrow(DT_model_sample) * 0.025

#usage parameters that control aspects of the rpart fit
my_control <- rpart.control(minsplit = min_node_split * 3,
              minbucket = min_node_split, 
              xval = 10,
              surrogatestyle = 0)

set.seed()
#when error message check for factor level - shouldn't be greater than 53
# Fit an initial rpart model
rpart_model_init <- rpart(target ~ .,
                          data = DT_model_sample[, .SD, .SDcols = analys_cols],
                          #data = DT[, -c("Id"),with = FALSE],
                          method = method_m,
                          control = my_control)

# display cp table -> smaller CP value indicates a simpler tree,  It represents the cost of adding another variable to the mode
printcp(rpart_model_init)

#plot cp value
plotcp(rpart_model_init)

#plot tree - use tweak to adjust font size. 1 is default
rpart.plot(rpart_model_init, type = 3, extra = 1, digits = 2,
           fallen.leaves = TRUE, roundint = TRUE, cex = NULL,
           tweak = 1.0, shadow.col = "grey")

#print confusion matrix
confusionMatrix(DT_model_sample$target, predict(rpart_model_init, newdata = DT_model_sample, type = method_m))

# show population of end nodes
print(rpart_model_init$frame$wt[rpart_model_init$frame$var == "<leaf>"])

rm(min_node_split, my_control, orig_cols, orig_num_cols)
```

## Prune the tree

```{r}
# get cp of lowest xstd - cross-validation error (xstd). Low xstd incidation of less variability in model's performance across cross-validation splits
cp_min_xstd <- rpart_model_init$cptable[which.min(rpart_model_init$cptable[, "xstd"]), "CP"]

# prune using cp_min_xstd
rpart_model <- prune(rpart_model_init, cp = cp_min_xstd, method_m = method_m)

rpart.plot(rpart_model, type = 3, extra = 1, digits = 2,
           fallen.leaves = TRUE, roundint = TRUE, cex = NULL,
           tweak = 1.0, shadow.col = "grey")

#show most important variables
varImp(rpart_model, drop = TRUE) 

#create predictions based on model
confusionMatrix(DT_model_sample$target, predict(rpart_model, newdata = DT_model_sample, type = method_m))

# show population of end nodes
print(rpart_model$frame$wt[rpart_model$frame$var == "<leaf>"])

rm(cp_min_xstd)
gc()
```

## Create chaid interaction variables

```{r include=FALSE}
# Predict the nodes using the party model
DT[, node_rank := rpart.predict.leaves(rpart_model, newdata = DT, type = "where")]

#check counts
freq(DT$node_rank)

# Extract the frame object from rpart_model
rpart_frame  <- data.table(rpart_model$frame)

# create row count
rpart_frame[, node_rank := as.integer(.I)]

#select end nodes - called <leaf>
rpart_frame <- rpart_frame[var == "<leaf>", .(node_rank, yval2.nodeprob)]

# join DT & rpart_frame
DT <- DT[rpart_frame, on = "node_rank"]

#convert node_score to char for binning
DT[, node_rank := as.character(node_rank)]

# needed for suffix assignment below
my_DT_cols = ncol(DT)

#create binaries from node score
DT <- dummy_cols(
  DT,
  select_columns = "node_rank",
  remove_first_dummy = FALSE,
  remove_most_frequent_dummy = FALSE,
  ignore_na = FALSE,
  split = NULL,
  remove_selected_columns = TRUE)  

# define the suffix
suffix = "_cbin"

# Get column names to append suffix
cols_to_update <- colnames(DT)[my_DT_cols:length(DT)]

# Append the suffix to the selected columns
setnames(DT, old = cols_to_update, new = paste0(cols_to_update, suffix))

rm(suffix, cols_to_update, my_DT_cols, rpart_frame)
gc()
```

# Identify corr, nzv and linear depend

Create analysis sample if records \> 50,000

```{r}
if (nrow(DT) > 100000) {
   DT_analysis_sample = DT[sample(.N, 50000)]
} else {
   DT_analysis_sample <- DT
}
```

Near zero variance

```{r}
nzv_cols <- nearZeroVar(DT_analysis_sample[, -c("target", "Id"), with = FALSE], 
                        names = TRUE,
                        saveMetrics = FALSE,
                        freqCut = 95/5,
                        uniqueCut = 10,
                        allowParallel = TRUE)

print(nzv_cols) 


```

High correlation variables

```{r include=FALSE}
# Get numeric column names & remove Id 
num_cols <- colnames(DT)[sapply(DT, is.numeric) & names(DT) != "Id"]  

# split work by 10, only upper triangle, use 64 bit algorithm
cor_matrix <- fastCor(DT_analysis_sample[, ..num_cols], nSplit = 10, upperTri = TRUE, optBLAS = TRUE)

# find the indices of the correlations above 0.8
highCorr_index <- which(cor_matrix > 0.8, arr.ind = TRUE)

# extract the pair names from the column names
highCorr_names <- paste(colnames(DT)[highCorr_index[, 1]], colnames(DT)[highCorr_index[, 2]], sep = " - ")

# print or save the pair names and the correlations
high_cor_names <- as.data.table(cbind(highCorr_names, cor_matrix[highCorr_index]))

#cleanup 
rm(cor_matrix, highCorr_index, high_cor_names)

```

Linear dependencies

```{r}
# find linear dependencies
lin_depend <- findLinearCombos(DT_analysis_sample[, ..num_cols])  

# extract names of column indexes
column_names <- lapply(lin_depend$linearCombos, function(index_list) {
  colnames(DT)[index_list]
})

# print names
print(column_names)
end_time <- Sys.time()
processing_time <- end_time - start_time
print(processing_time)
```

## Remove nzv variable

```{r}
#delete near zero variance columns 
#if (length(nzv_cols) > 0) {
#  DT[, (nzv_cols) := NULL]}
  # cleanup 
rm(nzv_cols, num_cols)



rm(DT_sample)
```
