---
title: "R Notebook"
author: "Tom Richebacher"
date: "7/11/2023"
output: html_notebook
---

# CALL LIBARIES {#sec-call-libaries}

```{r call-libraries, warning=TRUE, include=FALSE}

library(formatR)
library(compare)
library(randomForest)
library(HiClimR)
library(vtreat)
library(WVPlots)
library(wrapr)
library(lintr)
library(data.table)
library(caret)
library(ggplot2)
library(MLmetrics)
library(janitor)
library(partykit)
library(stringr)
library(DataExplorer)
library(smbinning)
library(SmartEDA)
library(fastDummies)
library(rattle)
library(rpart.plot)
library(RColorBrewer)
library(treeClust)
library(fst)
library(categoryEncodings)
library(corrplot)
library(doParallel)

detectCores()

cl <- makeCluster(8)

registerDoParallel(cl)
```

# [INITIALIZE GLOBAL VARIABLES]{.underline} {#sec-initialize-global-variables}

```{r initialize-global-variables, warning=TRUE}
options(scipen = 999)
use.integer64 <- FALSE

#"REG" = regression, "B_CLASS" = binary class, "M_CLASS" = multiclass
analysis_type <- "B_CLASS"
input_files <- 1
project <- "Credit"
report_path <- "C:/Users/trich/OneDrive/R_code/rtest/R_testing/reports/"
data_output_path <- "C:/Users/trich/OneDrive/R_code/rtest/R_testing/data_output/"
data_input_path <- "C:/Users/trich/OneDrive/R_code/rtest/R_testing/data_input/"
#data_input_path <- "C:/Users/trich/OneDrive/R_code/credit/"
set.seed(419)
```

# MY FUNCTIONS {#sec-my-functions}

```{r my-functions}
#**************************Fast Frequency counts***************************
# a little fast frequency count function
fast_freq <- function(data, column_names) {
  result <- lapply(column_names, function(column_name) {
    data[, .N, by = column_name]
  })
  return(result)
}
#***************************Perform KS statistics*********************************
#result_list <- fast_freq(DT, c("months_loan_duration"))
# Function to perform KS test
perform_ks_test <- function(col_name) {
  ks_result <- ks.test(DT_train_treated[[col_name]], 
                       DT_test_treated[[col_name]])
  return(ks_result)
}
#**************************char column to numeric var******************************
# Define a function to convert a char column to ordered numeric values
convert_to_ordered_numeric <- function(data, col_name, levels) {
  data[, (paste0(col_name, "_ord")) := as.numeric(factor(get(col_name), levels = levels, ordered = TRUE))]
}
#******************************Read text file**************************************
#*read_data <- function(filename) {
       fread(filename,
                # Set separator as comma
                sep = ",",
                # Read all rows - use nrows = xxx to read partial file
                nrows = -1,
                # Check column names for validity
                check.names = TRUE,
                #read random sample
                #n = 50000,
                # Set header as first row
                header = TRUE,
                # treat all integer columns as 64-bit integers
                #integer64 = 'numeric',
                #Use quote to specify character used to quote strings in the file
                #quote = "'",
                #specify the possible string representations of missing values
                na.strings = c("NA", "N/A", "null", "", "unknown"))
 }

```

# [GET FILES:]{.underline} {#sec-get-files}

## Read data {#sec-read-data}

```{r read-dataa, message=TRUE, warning=TRUE}

# Call function to read in data files
if (input_files == 1) {
  DT <- read_data(paste0(data_input_path, "DT.csv"))
  saveRDS(DT, paste0(data_input_path, "DT.Rds"))
} else if (input_files == 2) {
  train <- read_data("train.csv")
  train[, file := "train"]
  saveRDS(train, paste0(data_input_path, "train.Rds"))
  test <- read_data("test.csv")
  test[, file := "test"]
  saveRDS(test, paste0(data_input_path, "test.Rds"))
}

rm(read_data)




#for speed testind a data dublication function
#DT <- rbindlist(rep(list(DT), 10000))
```

## Stack train and test file {#sec-stack-train-and-test-file}

```{r eval=FALSE, include=FALSE}
if (input_files == 2) {
  # function to identify column header differences
  get_feature_names <- function(file1, file2) {
  # find the names that are common between the tables
  var_names <- intersect(names(file1), names(file2))
  # find columns different in file1
  var_names_file1 <- setdiff(names(file1), names(file2))
  # find columns different in file 2
  var_names_file2 <- setdiff(names(file2), names(file1))
  # show all differences between the two files
  col_diff <- paste0(var_names_file1, var_names_file2)
  return(col_diff)
}
  
  # return column header difference
  get_feature_names(train, test)

  # #assign Id if needed setnames(train, 'train_id', 'Id') setnames(test, 'test_id', 'Id')
  # stack train and test
    DT <- rbind(train, test, fill = TRUE, use.names = TRUE)
    rm(train, test, get_feature_names)
}

```

## Join Files

```{r eval=FALSE, include=FALSE}

DT <-  merge(DT, people, by = "people_id", all.x = TRUE)
DT$Id <- seq.int(nrow(DT))
 
rm(people)
```

## Create record Id

```{r}
#create index
DT[, Id := as.integer(.I)]

#move index to 1st column
setcolorder(DT, c("Id", colnames(DT)[1:(ncol(DT) - 1)]))
```

# [DATA SUMMARY REPORTS]{.underline}

Sample file if nrow \> 50,000

```{r}
if (nrow(DT) > 50000) {
   DT_sample_1 <- DT[sample(nrow(DT), 50000, replace = FALSE), ]
} else {
   DT_sample_1 <- DT
}
```

## Distribution Report

```{r paged.print=TRUE}
# Table with basic statistics
Distribution_report <- smbinning.eda(DT_sample_1,
                                     rounding = 2)$eda


gc()
```

# [COLUMN WORK]{.underline} - cleaning

## Column headers cleaning

```{r eval=FALSE, include=FALSE}
# identify non standard column headers
# Define a regex pattern for non-standard column names

non_standard_pattern <- "^[^A-Za-z_].*$"

# Identify non-standard column names
non_standard_cols <- names(DT)[grepl(non_standard_pattern, names(DT))]

# Print the non-standard column names
print(non_standard_cols)

# Clean column names
setnames(DT, clean_names(names(DT)))

setnames(DT, old = c("a", "d"), new = c("anew", "dnew"))

rm(non_standard_pattern, non_standard_cols)
```

## Columns class changes

```{r eval=FALSE, include=FALSE}
# If a column contains values that are strictly whole numbers (e.g., counts, IDs, years, months) make integer.

#individual var
DT[, target := as.factor(target)]

#by specific vars
change_col_class <- c("return_dollar", "product_id")

#by column text spring
change_col_class <- grep("(cat|bin)", names(DT), value = TRUE)

#by all vars of a specific class
change_col_class <- colnames(DT[, .SD, .SDcols = sapply(DT, is.numeric)])

#identify binaries to convert
change_col_class <- names(Filter(function(x)uniqueN(na.omit(x)) <= 2 & max(x) == 1, DT))

#********************************************************************************#
#do actual conversion
DT[, (change_col_class) := lapply(.SD, as.numeric), .SDcols = change_col_class]
#********************************************************************************#
#*
rm(change_col_class)
```

## Append identifier to original columns

```{r}
# get columns to exclude
exclude_columns <- c("default", "Id")

# Get original data columnns less those to exclude
orig_columns <- setdiff(colnames(DT), exclude_columns)

# define the suffix
suffix <- "_orig"

# Append the suffix to the selected columns
setnames(DT, old = orig_columns, new = paste0(orig_columns, suffix))

rm(orig_columns, exclude_columns, suffix)
```

# TARGET PREP

## Target Analysis & Prep

```{r}
# assign target name
setnames(DT, "default", "target")

# move target to 1st column
setcolorder(DT, c("target", setdiff(names(DT), "target")))

# target distribution
if (analysis_type != "REG") {
    DT[, .(.N), by = .(target)]
    plot_bar(DT$target)
} else {
    plot_histogram(DT$target)
}

# rename target value when binary
if (analysis_type == 'B_CLASS') {
  DT[, target := as.factor(target)]
  # Switch levels 2 to 1 and 1 to 0
  DT[, target := factor(fifelse(target == "2", "1", "0"))]
#DT[, target := ifelse(target == 1, "HaveNot", ifelse(target == 2, "Have", NA))]
}

# remove missing target records
DT[!is.na(target)]
gc()
```

## Normalize Target Variable

```{r eval=FALSE, include=FALSE}
DT[, target := log(target + 1)]
```

# CREATE BYPASS TABLE {#sec-create-bypass-table}

```{r include=FALSE}
# identify vars for bypass table
mov_var <- c("target", "Id")

# create bypass table
DT_bypass_file <- DT[, ..mov_var]

# Store the bypass table to disk in compressed version
write.fst(DT_bypass_file, paste0(data_output_path, "DT_bypass_file.fst"), 100)

# Retrieve the data 
# DT_report.Rds <- read.fst(paste0(data_path, "/DT_report.Rds"),
#          columns = NULL,
#          from = 1,
#          to = NULL,
#          as.data.table = FALSE,
#          old_format = FALSE)

# remove bypass table unless needed
rm(mov_var, DT_bypass_file)
gc()
```

# DATA EXPLORATION & REPORTING

## Data profile report

```{r}
# create in depth data report
Data_profile_report <- create_report(DT_sample_1, y = "target", report_title = paste0(project, "Data Profile Report"), output_file = paste0(project, "_Data_Profile Report"), output_dir = report_path)
```

## Categorical variable report

```{r include=FALSE}
# analysis of categorical data
Report_cat_vars <- ExpCatStat(DT_sample_1[, -c("Id")],Target = "target",
                                  result = "Stat",
                                  clim = 10,        #max cat levels allowed
                                  nlim = 10,        #max unique values for num vars
                                  bins = 10,        #max number of bins
                                  Pclass = "Have",
                                  plot = TRUE,
                                  top = 20,
                                  Round = 2)

#save report to disk
saveRDS(Report_cat_vars, file = paste0(report_path,
                                               file ="Report_Categorical_vars.rds"))
```

## Categorical detail vars to target report

```{r include=FALSE}
Report_cat_detail_vars<- ExpCatStat(DT_sample_1[, -c("Id")], Target = "target",
           result = "IV",
           clim = 5,          #max # of cats allowed
           nlim = 10,         # of levels for num vars
           bins = 10,         #max # of bins for cat & num vars
           Pclass = "Have")

saveRDS(Report_cat_detail_vars, file = paste0(report_path,                                               file ="Report_categorical_detail_vars.rds"))
```

## Numeric vars to target report

```{r include=FALSE}
Report_numerical_vars <- ExpNumStat(DT_sample_1[, -c("Id")],
                                    by = "G",
                                    gp = "target",
                                    Qnt = NULL,
                                    Nlim = 10,
                                    MesofShape = 2,
                                    Outlier = TRUE,
                                    round = 2,
                                    weight = NULL,
                                    dcast = FALSE,
                                    val = NULL)

saveRDS(Report_numerical_vars, file = paste0(report_path,                                               file ="Report_numerical_vars.rds"))
```

# [DATA CLEANING]{.underline} {#sec-basic-data-cleaning}

## Rows

### Row missing values

```{r eval=FALSE, include=FALSE}
# Calculate the number of missing values in each row

nrows = nrow(DT) #how many rows of data

rows_complete = sum(complete.cases(DT))  #how many complete rows

rows_complete / nrows  #calculate rows missing percentage

DT[, row_miss_orig := rowSums(is.na(DT))] #assign NA incidator in DT

table(DT$row_miss_orig)  #frequency count of missing by row

# filter out records if needed
#dt[row_miss_orig <= 0.30]
```

### Row deletes by column ( 'x' or 'y' have NA)

```{r eval=FALSE, include=FALSE}
# select columns
omit_row <- c("x", "y")

# see how many records are left
DT[complete.cases(DT[, ..omit_row]), ]
```

## Columns {#sec-columns}

### Unnecessary column removal {#sec-unnecessary-column-removal}

```{r eval=FALSE, include=FALSE}
#delete columns with one value
rm_cols <- colnames(DT)[sapply(DT, function(x) length(unique(x))) == 1]

#delete specific cols
rm_cols <- c("age_orig", "housing_orig")

if (length(rm_cols) > 0) {
  DT[, c(rm_cols) := NULL]
}

rm(rm_cols)
```

### Column missing value graph

```{r eval=FALSE, include=FALSE}
plot_missing(DT_sample_1) + theme_bw() + ggtitle("Percent Missing Values")  
```

### Impute incorrect NA values

```{r eval=FALSE, include=FALSE}

#Identify all columns with missing values
cols_with_NA <- names(DT)[colSums(is.na(DT)) > 0]

# replace some value in table like -1 with NA
replace_value <- -1

DT[, (cols_to_modify) := lapply(.SD, function(x) replace(x, x == replace_value, NA)), .SDcols = cols_to_modify]

rm(cols_with_NA, replace_value)
```

### Character to ordinal column manual recode

```{r include=FALSE}

a <- DT[, unique(savings_balance_orig)] #get the var data, exlude NAs

# Specify levels for each column
checking_levels <- c("< 0 DM", "1 - 200 DM", "> 200 DM")
savings_levels <- c("< 100 DM",  "101 - 500 DM", "501 - 1000 DM", "> 1000 DM")

# Apply the function for each column
convert_to_ordered_numeric(DT, "checking_balance_orig", checking_levels)
convert_to_ordered_numeric(DT, "savings_balance_orig", savings_levels)

rm(a, checking_levels, savings_levels)
```

# [ANALYTICAL COLUMN REMOVAL]{.underline}

## Remove high correlation vars

```{r eval=FALSE, include=FALSE}
# Get numeric column names & remove Id 
num_cols <- setdiff(colnames(DT_sample_1)[sapply(DT_sample_1, is.numeric)], "Id")

cor_matrix <- cor(DT_sample_1[, ..num_cols], use = "pairwise.complete.obs")

corrplot(cor_matrix, method = "color")

summary(cor_matrix[upper.tri(cor_matrix)])

highly_correlated_vars <- findCorrelation(cor_matrix, cutoff = 0.75, names = TRUE)

DT[, -highly_correlated_vars]

cor_matrix2 <- cor(DT_sample_1[, ..num_cols], use = "pairwise.complete.obs")

summary(cor_matrix[upper.tri(cor_matrix)])

#cleanup 
rm(cor_matrix, highly_correlated_vars)
```

## Remove near-zero-variance data

```{r eval=FALSE, include=FALSE}
nzv_cols <- nearZeroVar(DT_sample_1[, -c('target', 'Id')], 
                        names = TRUE,
                        saveMetrics = FALSE,   #make TRUE to run report
                        freqCut = 95/5,
                        uniqueCut = 10,
                        allowParallel = TRUE)

print(nzv_cols) 

#checkConditionalX(DT_sample_2, y2)
#delete near zero variance columns 
if (length(nzv_cols) > 0) {
  DT[, (nzv_cols) := NULL]
}

# cleanup 
rm(nzv_cols)
```

# [CHAID INTERACTION VARIABLES]{.underline}

```{r}
if (analysis_type == 'REG') {
  method_m = 'anova' 
  } else {
  method_m = 'class'
}

#minimum oberservation in a terminal node is # of records * x
min_node_split = nrow(DT_sample_1) * 0.025

#usage parameters that control aspects of the rpart fit
my_control <- rpart.control(minsplit = min_node_split * 3,
              minbucket = min_node_split, 
              xval = 10,
              surrogatestyle = 0)

#when error message check for factor level - shouldn't be greater than 53
# Fit an initial rpart model
rpart_model_init <- rpart(target ~ .,
                          data = DT_sample_1,
                          #data = DT[, -c("Id"),with = FALSE],
                          method = method_m,
                          control = my_control)

# display cp table -> smaller CP value indicates a simpler tree,  It represents the cost of adding another variable to the mode
printcp(rpart_model_init)

#plot cp value
plotcp(rpart_model_init)

#plot tree - use tweak to adjust font size. 1 is default
rpart.plot(rpart_model_init, type = 3, extra = 1, digits = 2,
           fallen.leaves = TRUE, roundint = TRUE, cex = NULL,
           tweak = 1.0, shadow.col = 'grey')

#print confusion matrix
confusionMatrix(DT_sample_1$target, predict(rpart_model_init, newdata = DT_sample_1, type = method_m))

# show population of end nodes
print(rpart_model_init$frame$wt[rpart_model_init$frame$var == '<leaf>'])

rm(min_node_split, my_control, orig_num_cols)
```

## Prune the tree

```{r}
# get cp of lowest xstd - cross-validation error (xstd). Low xstd incidation of less variability in model's performance across cross-validation splits
cp_min_xstd <- rpart_model_init$cptable[which.min(rpart_model_init$cptable[, "xstd"]), "CP"]

# prune using cp_min_xstd
rpart_model <- prune(rpart_model_init, cp = cp_min_xstd, method_m = method_m)

rpart.plot(rpart_model, type = 3, extra = 1, digits = 2,
           fallen.leaves = TRUE, roundint = TRUE, cex = NULL,
           tweak = 1.0, shadow.col = 'grey')

#show most important variables
varImp(rpart_model, drop = TRUE) 

#create predictions based on model
confusionMatrix(DT_sample_1$target, predict(rpart_model, newdata = DT_sample_1, type = method_m))

# show population of end nodes
print(rpart_model$frame$wt[rpart_model$frame$var == '<leaf>'])

rm(cp_min_xstd, DT_model_sample)
gc()
```

## Create interaction vars

```{r}
# Predict the nodes using the party model
DT[, node_rank := rpart.predict.leaves(rpart_model, newdata = DT, type = 'where')]

#check counts
table(DT$node_rank)

#convert node_score to char for binning
DT[, node_rank := as.character(node_rank)]

rm(rpart_model, rpart_model_init, DT_sample_1)
gc()
```

# [CHARACTER & FACTOR ENCODING]{.underline} {#sec-character--factor-encoding}

## Character-target encoding of nominal variables using vtreat

### Three-way sampling - with designtreatment (vtreat)

```{r eval=FALSE, include=FALSE}
# Create indices for a training set (60%), validation set (20%), and test set (20%)
set.seed(123)
calibration_indices <- createDataPartition(DT$target, p = 0.6, list = FALSE)
remaining_data <- DT[-train_indices, ]

validation_indices <- createDataPartition(remaining_data$target, p = 0.5, list = FALSE)
test_indices <- setdiff(seq_len(nrow(remaining_data)), validation_indices)

# use to create treatment plan
DT_calibration_data <- DT[calibration_indices, ]
DT_train_data <- remaining_data[validation_indices, ]
DT_test_data <- remaining_data[test_indices, ]


#remove target variable
all_columns <- setdiff(colnames(DT_calibration_data), "target")
# build the data treatments on calibration data

DT_calib_treatment_plan = designTreatmentsC(DT_calibration_data,
                                            all_columns,
                                            outcomename = 'target',
                                            outcometarget = '1',
                                            verbose = FALSE,
                                            missingness_imputation = NULL,
                                            minFraction = 0.02,
                                            parallelCluster = cl,
                                            use_parallel = TRUE)

DT_treatment_plan_detail <- setDT(DT_calib_treatment_plan$scoreFrame)

DT_train_treated <- prepare(DT_calib_treatment_plan, DT_train_data, pruneSig = NULL)
```

### Two-way sampling with crossframetreatment (vtreat)

```{r two-way-vtreat}
# Create the training and test sets

trainIndex <- createDataPartition(DT$target, p = 0.60, list = FALSE)
DT_train_data <- DT[trainIndex, ]
DT_test_data  <- DT[-trainIndex, ]

#save DT
write.fst(DT, paste0(data_output_path, "/DT.fst"), 100)
#remove DT
rm(DT)


if (analysis_type == 'B_CLASS'){
  vtreat_treatment <- 'mkCrossFrameCExperimnent'
} else if (analysis_type == 'N_CLASS') {
  vtreat_treatment <- 'mkCrossFrameNExperiment'
} else {
  vtreat_treatment <- 'mkCrossFrameMExperiment'
}

#create treatment
unpack[
  DT_train_treatment_plan = treatments,               #treatmen plan
  DT_train_treated = crossFrame                 #treated training data
] <- mkCrossFrameCExperiment(
  dframe = DT_train_data,
  varlist = setdiff(colnames(DT_train_data), "target"),
  outcomename = "target",
  outcometarget = 1,
  verbose = FALSE,
  missingness_imputation = -111,
  minFraction = 0.02,
  parallelCluster = cl,
  use_parallel = TRUE
  )

#convert scoreFrame to data table
DT_train_treatment_plan_detail <- setDT(DT_train_treatment_plan$scoreFrame)

#Recommendet variables
recommended_vars <- DT_train_treatment_plan_detail[recommended == TRUE, varName]
print(recommended_vars)

#apply treatment to test data - don't convert to data.table
DT_test_treated <- prepare(DT_train_treatment_plan, DT_test_data)

#apply suffix "treat" to train and test data
suffix <- "_treat"

# Get column names to append suffix
cols_to_update <- setdiff(colnames(DT_test_treated), c("target", "Id"))

for (cnt in list(DT_test_treated, DT_train_treated)) {
  setnames(cnt, old = cols_to_update, new = paste0(cols_to_update, suffix))
}
rm(trainIndex, suffix, cols_to_update, recommended_vars, DT_test_data, DT_train_data, cnt)
```

#### Treatment Analysis

```{r}

```

#### Analysis between train and test data

```{r warning=FALSE}

ks_columns <- setdiff(colnames(DT_train_treated), c("target", "Id"))

# call function "perform_ks_test"
ks_test_results <- lapply(ks_columns, perform_ks_test)

# Combine results into a data.table
ks_test_results_dt <- data.table(
  Column = ks_columns,
  KS_statistic = sapply(ks_test_results,
                        function(result) result$statistic),
                        p_value = sapply(ks_test_results,
                                         function(result) round(result$p.value,
                                          digits = 2)))


#convert treated dataset to data.table
setDT(DT_train_treated)
setDT(DT_test_treated)

#assign replacement value
replace_value <- -111

#columns to have replacement
cols_to_modify <- colnames(DT_test_treated)

#replace the vtreat assigned -111 with NA
for (file_name in list(DT_test_treated, DT_train_treated)) {
  file_name[, (cols_to_modify) := lapply(.SD, function(x) replace(x, x == replace_value, NA)), .SDcols = cols_to_modify]
  }

#save treatment plan
saveRDS(DT_train_treatment_plan, paste0(report_path, "DT_train_treatment_plan.Rds"))

#save treatment detail
saveRDS(DT_train_treatment_plan_detail, paste0(report_path, "DT_train_treatment_plan_detail.Rds"))

#save train_treated in binary file
write.fst(DT_train_treated, paste0(data_output_path, "DT_train_treated.fst"), 100)

#save test_treated in binary file
write.fst(DT_test_treated, paste0(data_output_path, "DT_test_treated.fst"), 100)

rm(ks_columns, ks_test_results, ks_test_results, p_value, DT_train_treated, file_name, replace_value, cols_to_modify)
```

K[S STATISTICS SIGNIFICANTS LEVEL]{.underline}

-   **0.00 to 0.20:** The two distributions are very similar.

-   **0.20 to 0.40:** The two distributions are somewhat different.

-   **0.40 to 0.60:** The two distributions are moderately different.

-   **0.60 to 0.80:** The two distributions are substantially different.

-   **0.80 to 1.00:** The two distributions are extremely different.

# [OTHER CODE]{.underline}

Row level - binary pattern creation

```{r eval=FALSE, include=FALSE}

#identify binary variables
create_bin_vars <- names(Filter(function(x) uniqueN(na.omit(x)) <= 2 & max(x) == 1, DT))

#create variable that sums the values in Bin_vars 
DT[, DT_sums  :=  rowSums(.SD, na.rm = TRUE), .SDcols = c(create_bin_vars)]

#compute mean of values in Bin_vars 
DT[, DT_means := rowMeans(.SD, na.rm = TRUE), .SDcols = c(create_bin_vars)]

#concatenate binary variable across columns and make it a factor variable
DT[, bin_pattern := do.call(paste0,.SD),.SDcols = create_bin_vars]
DT[, bin_pattern := as.factor(bin_pattern)]

# count zeros in row
DT[, num_cols_eq_0 := Reduce(sum, lapply(.SD, function(x) na.omit(x) == 0))]

#indicate if there are more than 4 missing values
DT[, high_nas := ifelse(amount_nas > 4, 1,0)] 

rm(Bin_vars)
```

Row level aggregation by specific columns

```{r eval=FALSE, include=FALSE}

#conditionalby header identification aggregation 
#product_up <- names(DT_tar)[which(regexpr("M_UP_1", names(DT_tar)) > 0)]
#DT_tar[, prod_up := rowSums(.SD), by=ncodpers, .SDcols = product_up]  

cols_title <-  c("reg", "car", "calc")
#cols_title <- names(DT3)[which(regexpr(p, names(DT3)) > 0)]

 for (p in cols_title) 
  {
    cols_title <- names(DT3)[which(regexpr(p, names(DT3)) > 0)]   
    DT3[, paste0(p, '_sum')    := rowSums(.SD, na.rm = TRUE),
        .SDcols = c(cols_title)] 
    DT3[, paste0(p, '_mean')   := rowMeans(.SD, na.rm = TRUE),
        .SDcols = c(cols_title)]
    DT3[, paste0(p, '_ratio')  := ifelse(get(paste0(p, '_mean')) == 0,0,
                                             (get(paste0(p, '_sum'))) / get(paste0(p, '_mean'))) ]
  }

```

Casting to wide data when multiple ID exist - like customer id

```{r eval=FALSE, include=FALSE}
#create new summary table with sum, mean, count


DT_summary_wide <- dcast(DT, customer_id ~ product_id + product_category,
                         value.var = c("revenue_dollar", "return_dollar"), 
                         fill = 0,
                         fun.aggregate = list(sum, mean, length))

#Create customer summary
DT_customer_summary <- DT[, .(dol_revenue = sum(revenue_dollar),
                              dol_returns = sum(return_dollar),
                              cnt_purchases = length(product_id),
                              cnt_prod_cats = .N,   
                              cnt_returns = .N),
                          by = .(customer_id)]

# Combine customer and order summaries
DT <- DT_summary_wide[DT_customer_summary, on = "Id"]

```

Column level

```{r eval=FALSE, include=FALSE}
#create new variable in DT
DT[,`:=`(u = z + 1, v = z - 1)] 
```

LAT/LONG WORK

```{r eval=FALSE, include=FALSE}
library(geosphere)

#Dist from centroid
DT$longmean <- mean((DT$longitude), na.rm = TRUE) / 1e6
DT$latmean  <- mean((DT$latitude), na.rm = TRUE) / 1e6

#Adjusted long lat
DT$longitude1 <- DT$longitude / 1e6
DT$latitude1 <- DT$latitude / 1e6

# Haversine distance
ncol(DT)
DT$geodist <- distHaversine(p[,31:32], DT[, 33:34])

rm_geo_cols <- c("longmean", "latmen", "longitude", "latitude1")

DT <- DT[(rm_geo_cols) := NULL]
```

Identify and collapse high cardinality vars

```{r}
{r eval=FALSE, include=FALSE}
library(Hmisc)
#get non-numeric columns, exclude target & require > 2 values
char_fact_cols <- colnames(DT)[!sapply(DT, function(x) {
  (is.numeric(x))}) & colnames(DT) != "target"]   


#test this
#target_vector <- DT$target
#lapply(DT_sample_2[, -c("Id")], table, target_vector)

# frequency distribution before combining
freq_before <- lapply(char_fact_cols, function(col) {
  freq_dt <- freq(DT_sample_2[, .(value = get(col))], plot = FALSE, na.rm = TRUE)
  freq_dt <- cbind(freq_dt, column = col)
  return(freq_dt)
})

# creates table that shows freqs before transformation
freq_before <- rbindlist(freq_before)

#each value in a var needs to have at least 5% coverage 
threshold = 0.05   

#combine levels
if (length(char_fact_cols) > 0) {
  DT[, (char_fact_cols) := lapply(.SD, function(x) combine.levels(x, minlev = threshold)), .SDcols = char_fact_cols]
}

# after combine frequency distributions
freq_after <- lapply(char_fact_cols, function(col) {
  freq_dt <- freq(DT_sample_2[, .(value = get(col))], plot = FALSE, na.rm = TRUE)
  freq_dt <- cbind(freq_dt, column = col)
  return(freq_dt)
})

# creates table with freqs after transformation
freq_after <- rbindlist(freq_after)

rm(threshold, char_fact_cols, DT_sample_2)
gc()
```

Create binary variable from reduced cardinality

```{r}
{r eval=FALSE, include=FALSE}
# get factor and char columns & exclude relevant columns
conv_to_bins <- colnames(DT)[sapply(DT, function(x) {
  (is.character(x) || is.factor(x)) && length(unique(x)) > 2
}) & colnames(DT) != "target"]

# length of the initial DT data 
DT_cols <- ncol(DT) + 1

if (length(conv_to_bins) > 0) {
DT <- dummy_cols(
  DT,
  select_columns = conv_to_bins,
  remove_first_dummy = TRUE,
  remove_most_frequent_dummy = FALSE,
  ignore_na = FALSE,
  remove_selected_columns = FALSE,
  split = NULL)}

# Clean column names
clean_names(DT)
# Replace symbols with their corresponding abbreviations

#replace < with LT
setnames(DT, colnames(DT), str_replace_all(colnames(DT), "<", "LT"))

#replace > with GT
setnames(DT, colnames(DT), str_replace_all(colnames(DT), ">", "GT"))

#replace <= with LE
setnames(DT, colnames(DT), str_replace_all(colnames(DT), "<=", "LE"))

#replace => with GE
setnames(DT, colnames(DT), str_replace_all(colnames(DT), "=>", "GE"))

# Replace parentheses with an underscore
setnames(DT, colnames(DT), str_replace_all(colnames(DT), "\\(|\\)", "_"))

# Remove empty spaces from column names
setnames(DT, colnames(DT), str_replace_all(colnames(DT), "\\s+", ""))

#replace varies symbols with underscore
setnames(DT, colnames(DT), str_replace_all(colnames(DT), "[&|/|,]", "_"))

# Remove underscores at the end of column names
setnames(DT, colnames(DT), str_replace_all(colnames(DT), "_$", ""))

# define the suffix to name the bin columns
suffix <- "_bin"

# Get the bin column names to update
cols_to_update <- colnames(DT)[DT_cols:length(DT)]

# Append the suffix to the selected columns
setnames(DT, old = cols_to_update, new = paste0(cols_to_update, suffix))

# remove unnecessary orig_ identifier in the bin column
setnames(DT, colnames(DT), str_replace_all(colnames(DT), "orig_", ""))

#clean-up
rm(conv_to_bins, DT_cols, cols_to_update, suffix)
gc()
```

Create weighted binaries

```{r eval=FALSE, include=FALSE}
# List of binary columns
weighted_bin_cols <- colnames(DT)[grep("_bin", colnames(DT))]

# get DT row count
DT_rows <- nrow(DT)


for (col in weighted_bin_cols) {
  new_col_name <- paste0(col, "_perc")
  set(DT, j = new_col_name, value = ifelse(DT[[col]] == 1, sum(DT[[col]] == 1) / DT_rows, 0))
}

rm(weighted_bin_cols, DT_rows)
gc()
```
